\newpage
\section{Matchings and Conectivity}
\lecture{9}{January 28, 2021}{YouTuhe, Lec. 8}

\begin{defn}[Matching]
	A \emph{matching} is a graph in which each vertex has degree at most $1$.

	A \emph{perfect matching} is a graph in which each vertex has degree $1$.
\end{defn}

\begin{thm}[Dilworth]
	Let $P$ be a partially ordered set with $n$ vertices. Then, there exists a chain of size $k$ (i.e., a sequence $v_1, v_2, \dots, v_k$ such that $v_i < v_j$ whenever $i < j$) or an anti-chain of size $n/k$ (i.e., a set of vertices such that for any pair $u, v$, $u \not< v$).
\end{thm}

\begin{thm}
	Let $G$ be a bipartite graph with $n$ vertices on each part. Let's call the parts $A$ and $B$.

	There exists a perfect matching inside $G$ if, and only if, for all subsets $S \subset A$, \[
		N(S) := \left|\cup_{u \in S}N(u)\right|	\ge |S|.
	\]
\end{thm}

\begin{sk}
	It is clear that it is a necessary condition. We shall prove that it is a sufficient condition. We'll use induction on $n$.

	Suppose that there is a set $A_1 \subset A$, $A_1 \neq A, \varnothing$ such that $|N(A_1)| = |A_1|$. Consider the graphs $G_1$ and $G_2$ by restraining the vertices to $A_1 \cup N(A_1)$ and $\overline{A_1 \cup N(A_1)}$, respectively. Show that $G_1$ and $G_2$ satisfy the hypothesis. Imply that there is a matching inside $G$.

	Suppose that $|N(S)| > |S|$ for all $S \subset A$, $S \neq A, \varnothing$. Pick pick any edge $uv$ and fix it. Consider  $G' = G - u - v$. Show that $G'$ satisfy the hypothesis. Imply that there is a matching inside $G$.
\end{sk}

\begin{prop} \label{prop:isolatedvertex}
	If $p < (1-\varepsilon)\frac{\log n}{n}$, then there exists an isolated vertex in $G(n, p)$ with high probability.
\end{prop}

\begin{dem} Let $X$ denote the number of isolated vertices in $G(n, p)$.
	\begin{align*}
		\Exp[X] &= n(1-p)^{n-1} \\
				&\gtrsim ne^{-(1-\varepsilon)\log{n}}\\
				&\gtrsim n^\varepsilon.
	\end{align*}

	On the other hand, 
	\begin{align*}
		\Exp[X^2] &= \sum_{u, v} \Prob(u, v \text{ are isolated})\\
				  &= n(n-1)(1-p)^{2n-3} + n(1-p)^{n-1}\\
				  &= \Exp[X]^2 \frac{n-1}{n}(1-p)^{-1} + \Exp[X].
	\end{align*}

	Therefore, 
	\begin{align*}
		\Var(X) &= \Exp[X^2] - \Exp[X]^2\\
				&\le 2p\Exp[X]^2 + \Exp[X]\\
				&\ll \EE[X]^2.
	\end{align*}

	Thus, by \nameref{lem:chebychevinq}, we're done.
\end{dem}

\begin{thm}
	Suppose $n$ is even.

	If $p < (1-\varepsilon)\frac{\log{n}}{n}$, then there is no perfect matching in $G(n, p)$ with high probability.

	If $p > (2+\varepsilon)\frac{\log{n}}{n}$, then there is a perfect matching in $G(n, p)$ with high probability.
\end{thm}

